---
title: "Visualizing surface displacements from BUnwarpJ"
author: "Jason Goetz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing surface displacements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
comment = "#>")
knitr::opts_knit$set(root.dir = "/Users/gnj/Dropbox/R-Projects/map.displ.r/inst/extdata/")
```

The [BUnwarpJ plugin](https://imagej.net/BUnwarpJ) in the open-source image processing package [ImageJ/Fiji](https://imagej.net/Fiji) can be a very useful tool for high-resolution surface deformation modelling from digital elevation models (DEMs). The `map.disp.r` package was built for importing bUnwarpJ image registration based on DEMs and computing displacements in the x,y and z directions. 

This vignette

* Describes and illustrates how to import raw transformation files exported from BUnwarpJ
* Shows how to create a raster of the displacements
* Shows how to visualize displacements directions and magnitudes


As a case study, we will map the deformation patterns of rock glaicer using DEMs that were derived from lidar and structure-from-motion topographic surveying [(Goetz et al 2019)](https://www.sciencedirect.com/science/article/pii/S0034425719302949).

### Data
To start, we will load our DEMs for the years 2012 and 2017 that were used for performing image registration in BUnwarpJ for displcament mapping.

```{r}
library(raster)
library(sp)

dem2012 <- raster("rock_glacier_dem_2012_50cm.tif")
dem2017 <- raster("rock_glacier_dem_2017_50cm.tif")
```

### Importing BUnwarpJ transformation file

The raw transformation file from BUnwarpJ describes the registration by providing the coordinates for each pixel of the source image and the corresponding row and column position of the target image. To apply this transformation to georeferenced images, we need to convert these positions to a local CRS. The `dem.displacement.mapping()` function does this for us by using the source and target DEMs used for performing image registration. Additionally, it computes the vertical displacement (in the z direction), as well as the direction (aspect) of the 2D (xy) displacements and the slope of the z displacement.

```{r}
library(map.displ.r)
f_tx <- "RAW_rock_glacier_hillshade_2012_50cm_direct_transf.txt"
d_tx <- dem.displacement.mapping(tx_file = f_tx, r_source = dem2012, r_target = dem2017)
head(d_tx)
```

Using `raster` we can assign the values of the displacement mapping to a grid. This is useful for exporting. 

```{r,  fig.height=6, fig.width = 6}

# Assign displacement values to an emptry raster to export
r_na <- setValues(dem2012, NA)
disp_3d <- setValues(r_na, d_tx$xyz_disp)
plot(disp_3d)
```

### Visualizing surface displacements

For visualization of the displacements in R, we can use `ggplot2` combined with the `metR` and `ggnewscale` packages. `metR` can create arrows indicating the magnitude and direction of displacement vectors, and `ggnewscale` lets us plot multiple `geom_raster()`'s each with their own fill scale and legend.

```{r,  fig.height=6, fig.width = 6}
library(ggplot2)
library(metR)
library(ggnewscale)

# Load hillshade for our map of displacements
hs2017 <- raster("rock_glacier_hillshade_2017_50cm.tif")
hs2017_df <- as.data.frame(hs2017, xy=TRUE)
names(hs2017_df) <- c("x", "y", "hs")

# Estimate mean annual surface velocity (m/yr) over the five year period
d_tx$mean_disp <- d_tx$xyz_disp/5

# Make displacement map
map <- ggplot(d_tx, aes(x_source,y_source)) +
  geom_raster(data=hs2017_df, aes(x=x, y=y, fill = hs), show.legend = FALSE) +
  scale_fill_gradient(high = "white", low = "black", na.value = "#FFFFFF") + 
  # Allow for multiple scale fills using ggnewscale package
  new_scale("fill") +
  
  geom_raster(data=d_tx, alpha = 0.4, aes(x=x_source, y=y_source, fill = mean_disp)) +
  scale_fill_viridis_c(name = "Mean annual\nsurface velocity\n(m/yr)", direction = 1) +
  
  # Create arrows using metR package
  geom_arrow(data=d_tx, aes(dx = x_disp, dy = y_disp), skip = 30, show.legend = FALSE) +
  
  xlab("Easting (m)") +
  ylab("Northing (m)") +
  coord_fixed() +
  theme(text = element_text(size = 9), axis.title = element_text(size = 9),
        axis.text = element_text(size = 6), axis.text.y = element_text(angle = 90))

map
```

### References

Arganda-Carreras, I., Sorzano, C. O., Marabini, R., Carazo, J. M., Ortiz-de-Solorzano, C., & Kybic, J. (2006, May). Consistent and elastic registration of histological sections using vector-spline regularization. In International Workshop on Computer Vision Approaches to Medical Image Analysis (pp. 85-95). Springer, Berlin, Heidelberg.

Goetz, J., Fieguth, P., Kasiri, K., Bodin, X., Marcer, M., & Brenning, A. (2019). Accounting for permafrost creep 
in high-resolution snow depth mapping by modelling sub-snow ground deformation. Remote Sensing of Environment, 231, 111275.

